To Airbnb?: A Question of Returns

### 2018-05-22 papers/full_paper.Rmd
Begin conversion of paper

---
title: "To Airbnb?: A Question of Returns"
subtitle: "Andy Krause and Gideon Aschwanden"
author: "Greenfield Advisors and University of Melbourne"
date: "October 25th, 2017"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
---

@@

  library(spdep)
  library(maptools)
  library(gstat)
  library(ggplot2)
  library(geosphere)
  library(ggmap)
  library(xtable)
  library(chron)
  library(plyr)
  library(sp)
  library(rgeos)
  library(reshape2)
  library(stringr)
  library(RColorBrewer)
  library(Hmisc)
  library(kernlab)
  library(lmtest)
  library(ROCR)

@@

```{r set_paths, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, results='hide'}

 data.path <- 'c:/dropbox/research/airBNB/data/'
 code.path <- 'c:/code/'
 fig.path <- 'c:/dropbox/research/airBNB/figures/'

 ## Set options

  options(xtable.comment = FALSE)
  options(xtable.caption.placement='top')

```{r source_files, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, results='hide'}

  source(paste0(code.path, "tools/dataviztools/ggplottools.R"))
  source(paste0(code.path, "tools/datamgmttools/dataMungetools.R"))
  source(paste0(code.path, "research/AirbnbMelbourne/abb_Functions.R"))

```{r set_colors, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE}

 # Set all colors

  abb.colors.names=c('rausch', 'hackberry', 'kazan', 'babu', 'lima', 'beach',
                     'ebisu', 'tirol', 'foggy', 'hof')
  abb.col <- c(rgb(250,88,99, maxColorValue=255), rgb(139,14,82, maxColorValue=255),
               rgb(0,117,140, maxColorValue=255), rgb(4,211,191, maxColorValue=255), 
               rgb(77,226,110, maxColorValue=255), rgb(252,179,14,
                                                       maxColorValue=255),
               rgb(249,177,139, maxColorValue=255), rgb(191,166,100,
                                                        maxColorValue=255),
               rgb(156,161,155, maxColorValue=255), rgb(86,94,97, maxColorValue=255))

  # Set specific colors

  str.col <- abb.col[1]
  ltr.col <- abb.col[3]

  strp.col <- abb.col[3]
  ltrp.col <- abb.col[6]

```{r load_workspace, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, cache=FALSE}

   load(paste0(data.path, '/analyzed/analyzed_workspace_sep.RData'))
   a<-1

@@

```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE}

  exch.rate <- 1.32

@@

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an Airbnb-type, short-term rental approach to generating income. To determine the likely LTR rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

@@

1. House:  Includes properties labeled 'House' or 'Townhouse'
2. Apartment: Includes properties labeled as 'Apartment' or 'Condominium'
3. Other: Includes properties labeled as 'Bed & Breakfast', 'Boat', 'Bungalow', 'Cabin', 'Camper/RV', 'Chalet', 'Dorm', 'Earth House', 'Hut', 'Igloo', 'Loft', 'Other', 'Tent', 'Treehouse', Villa' and 'Yurt.

@@

1. House:  Includes properties labeled 'Duplex', 'House', 'Terrace' or 'Townhouse'
2. Apartment: Includes properties labeled as 'Unit' or 'Studio'
3. Other: Includes properties labeled as 'Semi' and 'Villa'

@@

The Other category only makes up a very small percentage of the properties in both the short and long-term market; especially so in the long-term market (\autoref{typecomp}) .  Additionally, a large proportion of the 'Others' in the short-term data are Bed & Breakfast units, a use which doesn't fit within our research question. As a result, we remove the 'Others' from both the short- and the long-term rental data.

```{r plot_pt, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, results='hide', fig.height=3, fig.cap='\\label{typecomp}Property Type by Tenure'}

  load(paste0(fig.path, 'ltrtypes.RData'))
  load(paste0(fig.path, 'strtypes.RData'))

  str.types <- str.types + xlab('')
  ltr.types <- ltr.types + xlab('')

  ggMultiPlots(str.types, ltr.types, cols=2)

@@

As long-term rentals do not offer Private Room or Shared Room options^[Our long-term rental dataset does not include rooming houses, purpose built student accommodations and other types of rentals which may offer private or shared room options] and our purpose here is a comparison of long- and short-term returns, we remove all short-term properties that do not lease the entire home or apartment. Unfortunately, this filter does remove about 40% of the short-term data (\autoref{ltypecomp}), however, given the research question it is unavoidable.

```{r plot_ltype, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=2.5, fig.width=3.75, fig.cap='\\label{ltypecomp}Listing Type of Short-Term Rentals'}

 load(paste0(fig.path, 'listtypes.RData'))

 list.types + ggtitle('')

@@

In addition to looking at bedrooms and bathrooms as separate dimensions of properties, they can also be considered together.  In other words, it is very common in the industry (and when looking for a short-term rental) to specify searches by bed/bath combination for obvious reasons -- 1 bedroom with 3 baths is inefficient and 4 bedrooms with 1 bath is uncomfortable.  As a result we examine the combinations of the two (\autoref{bbtype}). From this analysis we see that there are 6 combinations that make up most all properties:  1bed/1bath, 2/1, 2/2, 3/1, 3/2 and 4/2. We filter the data to these combinations.

```{r plot_bb, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=4, fig.width=5.75, fig.cap='\\label{bbtype}Bed and Baths by Tenure by Type'}

 load(paste0(fig.path, 'ltrbb.RData'))
 load(paste0(fig.path, 'strbb.RData'))

 stbb <- stbb + ylab('# of Obs') + ggtitle('Short-Term Rentals')
 ltbb <- ltbb + ylab('# of Obs') + ggtitle('Long-Term Rentals')

 ggMultiPlots(stbb, ltbb)

@@

```{r plot_ratesrents, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=3, fig.width=5.75, fig.cap='\\label{rrdist}Distribution of Rates and Rents'}

 load(paste0(fig.path, 'rateplot.RData'))
 load(paste0(fig.path, 'rentplot.RData'))

 ggMultiPlots(rate.plot, ltr.plot, cols=2)

@@

Location is an important determinant in both the short- and long-term rental markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs. Inner suburbs are < 10km from the CBD, Middle suburbs 10 to 20km from CBD and Outer suburbs > 20km from CBD. In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high-end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts. The specific suburb of a property is the second, or finer scale, at which the market operates.

@@

<!--  loc.map <- loc.map + -->
<!--    theme(axis.title.y=element_blank(), -->
<!--          axis.text.y=element_blank(), -->
<!--          axis.ticks.y=element_blank(), -->
<!--          axis.title.x=element_blank(), -->
<!--          axis.text.x=element_blank(), -->
<!--          axis.ticks.x=element_blank()) -->

<!--  loc.map -->

@@

It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analyzing short and long-term rentals in Melbourne:

1. Rural (Outer suburbs, not Beach)
2. Suburban (Middle suburbs, not Beach)
3. City (Inner Suburbs, not Beach, not Core)
4. City-Core (Select inner suburbs with tourist activities and near CBD)
5. Beach (Properties within 500m of Port Phillip Bay east of Yarra River)

We select the 18 suburbs below to represent the 'core' of the city (Table 1).  The majority of tourist destinations and major events such as the Australian Open and the Grand Prix are located in these suburbs.  All are well served by public transportation and possess abundant amenities for tourists.

```{r table_corelist, results="asis", comment=FALSE, message=FALSE, echo=FALSE}

 c.subs <- data.frame(A=core.suburbs[1:6],
                      B=core.suburbs[7:12],
                      C=core.suburbs[13:18])
 names(c.subs) <- NULL
 rownames(c.subs) <- NULL

 print(xtable(c.subs, caption = 'List of Core Suburbs\n'),
       include.rownames=F,
       caption.placement="top",
       hline.after=c(0,6))

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs). The location of the short- and long-term rentals, colored by submarket, are shown in \autoref{pointmaps}.

Before assigning submarkets to the long-term property we must remedy the fact that about 5% of the long-term rental observations have missing latitude and longitude values.  Half of these do, however, have a latitude and longitude values for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps we apply the street centroid value to the observations with missing lat/long values.  For those without either set of lat/longs we remove them from the dataset at this point. After making this correction we then assign one of the five submarket designations based on the process described above.

```{r load_locations, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE}

  load(paste0(fig.path, 'ltrMap.RData'))
  load(paste0(fig.path, 'strMap.RData'))

```{r plot_locations, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=5, fig.height=10, fig.cap='\\label{pointmaps}Location Maps', fig.ext='png'}

  str.map <- str.map +
   theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
         axis.ticks.y=element_blank(),
         axis.title.x=element_blank(),
         axis.text.x=element_blank(),
         axis.ticks.x=element_blank())

  ltr.map <- ltr.map +
   theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
         axis.ticks.y=element_blank(),
         axis.title.x=element_blank(),
         axis.text.x=element_blank(),
         axis.ticks.x=element_blank())

  # ggMultiPlots(str.map, ltr.map, cols=1)

 To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (\autoref{smcounts}). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

```{r plot_locbars, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=8.5, fig.height=4, fig.cap='\\label{smcounts}Submarket Observation Counts'}

  load(paste0(fig.path, 'ltrsubplot.RData'))
  load(paste0(fig.path, 'strsubplot.RData'))

  ltrtype.map <- ltrtype.map + ylab("") + ggtitle('Long-Term Properties')
  strtype.map <- strtype.map + ggtitle('Short-Term Properties')

  ggMultiPlots(strtype.map, ltrtype.map, cols=2)

### Summary of Data Cleaning

We have applied ten separate filters to our two datasets. In the name of transparency and to highlight the bounds of our research question it is useful to review these filters (Table 2). There are more than 20,000 properties that are or have been listed on the Airbnb portal in Melbourne during the 2014 to 2016 period.  Filtering those not listed during the study period removes nearly 1/3 of all properties.  Further removing by property type and listing type (not entire home) elimates another 50% of the remaining properties.  From these 7,200 properties structural characteristics filters remove another 480 from consideration.  Over 1,000 of the remaining properties were never been booked and/or had outlying nightly rates and were removed.  Finally, 14 properties had either bad location data or missing information host policy. From over 20,000, we have 5,788 Airbnb properties meet the requirements of our research question.

For the long-term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 12,000 observations, leaving 120,092 in the final dataset.

```{r table_cleansummary, results="asis", comment=FALSE, message=FALSE, echo=FALSE}

 ## Extract data

  cc <- clean.count[,c('operation', 'str', 'ltr')]

 ## Sum counts and create percentages

  str.count <- cumsum(c(cc$str[1], -cc$str[-1]))
  ltr.count <- cumsum(c(cc$ltr[1], -cc$ltr[-1]))
  str.cut <- c(0, str.count[-length(str.count)] - str.count[-1])
  str.per <- c(0, str.cut[-1] / str.count[-length(str.count)])
  ltr.cut <- c(0, ltr.count[-length(ltr.count)] - ltr.count[-1])
  ltr.per <- c(0, ltr.cut[-1] / ltr.count[-length(ltr.count)])

 ## Compile into data.frame

  clean.table <- data.frame(field=cc$operation,
                          strcount=str.count,
                          strcut=str.cut,
                          strper=str.per,
                          ltrcount=ltr.count,
                          ltrcut=ltr.cut,
                          ltrper=ltr.per)

  ## Fix fields and names

  rownames(clean.table) <- NULL
  clean.table$field <- c('Initial Count', 'Time', 'Property Type',
                         'Listing Type', 'Bedrooms', 'Baths', 'Bed & Bath Combo',
                         'Rates & Rents', 'Revenue', 'Location', 'Host Policy')
  clean.table$strper <- 100*round(clean.table$strper, 2)
  clean.table$ltrper <- 100*round(clean.table$ltrper, 2)
  names(clean.table) <- c('Filter', 'Short-term', 'Cut', '% Cut',
                          'Long-term', 'Cut', '% Cut')

 ## Conver to xtable and print

  options(xtable.comment = FALSE)
  print(xtable(clean.table, 
               digits=0,
               caption='Data Cleaning Summary'),
       include.rownames=FALSE,
       caption.placement='top')

@@

1. Observed ($REV_o$)}: Observed bookings multiplied by median nightly rate
2. Actual ($REV_a$)}: Observed bookings extrapolated to a full year representing the likely actual revenue if the property had been listed all year long.^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] based on observed occupancy rate multiplied by median nightly rate
3. Potential ($REV_p$)}: Observed occupancy rate of non-blocked days multiplied by 366 multiplied by median nightly rate

As the observed calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the Actual revenue metric to be a better measure of actual market performance. We chart the distribution of the Actual ($REV_a$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties. Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year (\autoref{revdens}).  Actual short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not.  When looking at potential revenues, however, a great deal more may have the potential to produce additional revenue.  The high number of very low yield Airbnb properties in the Actual condition suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.]

```{r plot_revdens, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=6,fig.height=3.5, fig.cap='\\label{revdens}Comparison of Revenue Distributions'}

  ggplot(revdens.data, aes(x=revenue, fill=tenure, color=tenure)) +
    geom_density(alpha=.5) +
    scale_fill_manual(values=c(ltr.col, abb.col[2], str.col)) +
    scale_color_manual(values=c(ltr.col, abb.col[2], str.col)) +
    xlab('\nAnnual Revenue') +
    scale_x_continuous(breaks=c(seq(0,60000, by=20000)),
                       labels=c('$0', '$20k', '$40k', '$60k'))+
    theme(legend.position='none',
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y=element_blank()) +
    coord_cartesian(xlim=c(0, 70000)) +
    facet_wrap(~tenure)

In this section we compare net revenues between long- and short-term rental options.  As mentioned above, there are no direct matches between the datasets. In other words, we have not directly observed a property which has been in a long-term rental followed by short-term rental or vice versa.  We will, instead, make the comparisons between the two by imputing likely long-term rental revenues for the observed short-term properties and then comparing the measured net revenues of the short-term properties to their estimated imputed long-term revenues. We first focus on using the Actual measure of net revenue for the short-term properties as discussed above.  We save discussion and analysis of the Potential revenue for the sensitivity tests later in the study.

We make our imputations via hedonic price models. We model the long-term market with the long-term data and then use this model to predict or impute the likely long-term weekly rental prices for the short-term properties. We use the following model specification:

\begin{equation}
  \label{eq:test}
ln(RentLT_{i}) = \beta_{1} BedBath_{i} + \gamma_{2} Suburb_{i} + \tau_{3} Time_{i} + \epsilon_{i}
\end{equation}

where $ln(RentLT_{i})$ is the natural log of the weekly rent, *BedBath* is the bedroom/bathroom combination, *Suburb* is a fixed effect for the location of the property, *Time* is the month of rental and $\epsilon$ a random error term.

Realizing that the imputed long-term rental values may be more accurate when estimated at smaller scales, we impute long-term rental rates at submarket level, specifically by type, bed/bath and geographic area. In other words, we separately estimate LTR rents for Apartments with 1bed/1bath in the city core, Apartments with 1bed/1bath in the beach submarket, etc. for all combinations of type, bed/bath and geographic area. We then combine the results from all estimates back into a single dataset that contains all of the short-term rentals.

## Results

>>>>

---
title: "To Airbnb?: A Question of Returns"
subtitle: "Andy Krause and Gideon Aschwanden"
author: "Greenfield Advisors and University of Melbourne"
date: "May 5th, 2018"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
---

@@

  library(tidyverse)
  library(sf)
  library(geosphere)
  library(ggmap)
  library(multidplyr)
  library(tmap)

  options(stringsAsFactors=FALSE)
  options(xtable.comment = FALSE)
  options(xtable.caption.placement='top')

  data_path <- file.path('c:', 'dropbox', 'research', 'airBNB', 'data')
  code_path <- file.path('c:', 'code', 'research', 'airbnbmelbourne')
  fig_path <- file.path('c:', 'dropbox', 'research', 'airBNB', 'figures')

  source(file.path(code_path, 'functions', 'abb_Functions.R'))

@@

```{r load_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, results='hide'}

  load(file.path(data_path, 'analyzed', 'analyzed_data.RData'))

@@

@@

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental (LTR) versus an Airbnb-type, short-term rental (STR) approach to generating income. To determine the likely LTR rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

@@

1. House:  Includes properties labeled 'House', 'Townhouse' or 'Villa'
2. Apartment: Includes properties labeled as 'Apartment', 'Condominium' or 'Loft'
3. Other: Includes properties labeled as 'Bed & Breakfast', 'Boat', 'Bungalow', 'Cabin', 'Camper/RV', 'Chalet', 'Dorm', 'Earth House', 'Hut', 'Igloo', 'Other', 'Tent', 'Treehouse' and 'Yurt.

@@

1. House:  Includes properties labeled 'Duplex', 'House', 'Terrace', 'Townhouse' or 'Villa'
2. Apartment: Includes properties labeled as 'Unit' or 'Studio'
3. Other: Includes properties labeled as 'Semi'

@@

The Other category only makes up a small percentage of the properties in both the short (3.78%) and long-term (0.02%) markets.  Additionally, a large proportion of the 'Others' in the short-term data are Bed & Breakfast units, a use which doesn't fit within our research question. As a result, we remove the 'Others' from both the short- and the long-term rental data.

@@

As long-term rentals do not offer Private Room or Shared Room options^[Our long-term rental dataset does not include rooming houses, purpose built student accommodations and other types of rentals which may offer private or shared room options] and our purpose here is a comparison of long- and short-term returns, we remove all short-term properties that do not lease the entire home or apartment. Unfortunately, this filter does remove about 46% of the short-term data, however, given the research question it is unavoidable.

@@

In addition to looking at bedrooms and bathrooms as separate dimensions of properties, they can also be considered together.  In other words, it is very common in the industry (and when looking for a short-term rental) to specify searches by bed/bath combination for obvious reasons -- 1 bedroom with 3 baths is inefficient and 4 bedrooms with 1 bath is uncomfortable.  As a result we examine the combinations of the two. From this analysis we see that there are 6 combinations that make up most all properties:  1bed/1bath, 2/1, 2/2, 3/1, 3/2, 3/3, 4/2 and 4/3.  The four bedrooms options are only present in the house category. We filter the data to these combinations.

@@

..HOLD FOR DENSITY OF RATE AND RENTS PLOT..

<!-- ```{r plot_ratesrents, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=3, fig.width=5.75, fig.cap='\\label{rrdist}Distribution of Rates and Rents'} -->

<!--  load(paste0(fig.path, 'rateplot.RData')) -->
<!--  load(paste0(fig.path, 'rentplot.RData')) -->

<!--  ggMultiPlots(rate.plot, ltr.plot, cols=2) -->

@@

<!-- Location is an important determinant in both the short- and long-term rental markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs. Inner suburbs are < 10km from the CBD, Middle suburbs 10 to 20km from CBD and Outer suburbs > 20km from CBD. In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high-end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts. The specific suburb of a property is the second, or finer scale, at which the market operates.  -->

@@

<!--  loc.map <- loc.map + -->
<!--    theme(axis.title.y=element_blank(), -->
<!--          axis.text.y=element_blank(), -->
<!--          axis.ticks.y=element_blank(), -->
<!--          axis.title.x=element_blank(), -->
<!--          axis.text.x=element_blank(), -->
<!--          axis.ticks.x=element_blank()) -->
<!--   ggsave(loc.map, filename=file.path(fig.path, 'loc_map.png'), -->
<!--          bg='transparent', -->
<!--          width=6, height=6) -->
<!--  # loc.map -->

@@

<!-- ![Suburb Categories](c://dropbox/research/airBnb/figures/loc_map.png) -->

<!-- It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analyzing short and long-term rentals in Melbourne:  -->

<!-- 1. Rural (Outer suburbs, not Beach) -->
<!-- 2. Suburban (Middle suburbs, not Beach) -->
<!-- 3. City (Inner Suburbs, not Beach, not Core) -->
<!-- 4. City-Core (Select inner suburbs with tourist activities and near CBD) -->
<!-- 5. Beach (Properties within 500m of Port Phillip Bay east of Yarra River) -->

<!-- We select the 18 suburbs below to represent the 'core' of the city (Table 1).  The majority of tourist destinations and major events such as the Australian Open and the Grand Prix are located in these suburbs.  All are well served by public transportation and possess abundant amenities for tourists. -->

<!-- ```{r table_corelist, results="asis", comment=FALSE, message=FALSE, echo=FALSE} -->

<!--  c.subs <- data.frame(A=core.suburbs[1:6], -->
<!--                       B=core.suburbs[7:12], -->
<!--                       C=core.suburbs[13:18]) -->
<!--  names(c.subs) <- NULL -->
<!--  rownames(c.subs) <- NULL -->

<!--  print(xtable(c.subs, caption = 'List of Core Suburbs\n'), -->
<!--        include.rownames=F, -->
<!--        caption.placement="top", -->
<!--        hline.after=c(0,6)) -->

<!-- We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs). The location of the short- and long-term rentals, colored by submarket, are shown in \autoref{pointmaps}. -->

<!-- Before assigning submarkets to the long-term property we must remedy the fact that about 5% of the long-term rental observations have missing latitude and longitude values.  Half of these do, however, have a latitude and longitude values for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps we apply the street centroid value to the observations with missing lat/long values.  For those without either set of lat/longs we remove them from the dataset at this point. After making this correction we then assign one of the five submarket designations based on the process described above. -->

<!-- ```{r load_locations, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE} -->

<!--   load(paste0(fig.path, 'ltrMap.RData')) -->
<!--   load(paste0(fig.path, 'strMap.RData')) -->

<!-- ```{r plot_locations, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=5, fig.height=10, fig.cap='\\label{pointmaps}Location Maps', fig.ext='png'} -->

<!--   str.map <- str.map + -->
<!--    theme(axis.title.y=element_blank(), -->
<!--          axis.text.y=element_blank(), -->
<!--          axis.ticks.y=element_blank(), -->
<!--          axis.title.x=element_blank(), -->
<!--          axis.text.x=element_blank(), -->
<!--          axis.ticks.x=element_blank()) -->

<!--   ltr.map <- ltr.map + -->
<!--    theme(axis.title.y=element_blank(), -->
<!--          axis.text.y=element_blank(), -->
<!--          axis.ticks.y=element_blank(), -->
<!--          axis.title.x=element_blank(), -->
<!--          axis.text.x=element_blank(), -->
<!--          axis.ticks.x=element_blank()) -->

<!--     ggsave(str.map, filename=file.path(fig.path, 'str_map.png'), -->
<!--          bg='transparent', -->
<!--          width=6, height=6) -->
<!--     ggsave(ltr.map, filename=file.path(fig.path, 'ltr_map.png'), -->
<!--          bg='transparent', -->
<!--          width=6, height=6) -->

<!--   # ggMultiPlots(str.map, ltr.map, cols=1) -->

<!-- ![Short Term Rental Locations](c://dropbox/research/airBnb/figures/str_map.png) -->

<!-- ![Long Term Rental Locations](c://dropbox/research/airBnb/figures/ltr_map.png) -->

<!--  To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (\autoref{smcounts}). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size. -->

<!-- ```{r plot_locbars, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=8.5, fig.height=4, fig.cap='\\label{smcounts}Submarket Observation Counts'} -->

<!--   load(paste0(fig.path, 'ltrsubplot.RData')) -->
<!--   load(paste0(fig.path, 'strsubplot.RData')) -->

<!--   ltrtype.map <- ltrtype.map + ylab("") + ggtitle('Long-Term Properties') -->
<!--   strtype.map <- strtype.map + ggtitle('Short-Term Properties') -->

<!--   ggMultiPlots(strtype.map, ltrtype.map, cols=2) -->

<!-- ### Summary of Data Cleaning -->

<!-- We have applied ten separate filters to our two datasets. In the name of transparency and to highlight the bounds of our research question it is useful to review these filters (Table 2). There are more than 20,000 properties that are or have been listed on the Airbnb portal in Melbourne during the 2014 to 2016 period.  Filtering those not listed during the study period removes nearly 1/3 of all properties.  Further removing by property type and listing type (not entire home) elimates another 50% of the remaining properties.  From these 7,200 properties structural characteristics filters remove another 480 from consideration.  Over 1,000 of the remaining properties were never been booked and/or had outlying nightly rates and were removed.  Finally, 14 properties had either bad location data or missing information host policy. From over 20,000, we have 5,788 Airbnb properties meet the requirements of our research question. -->

<!-- For the long-term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 12,000 observations, leaving 120,092 in the final dataset. -->

@@

1. Actual ($REV_o$)}: Observed bookings multiplied by median nightly rate
2. Likely ($REV_l$)}: Observed bookings extrapolated to a full year representing the likely actual revenue if the property had been listed all year long.^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] based on observed occupancy rate multiplied by median nightly rate
3. Potential ($REV_p$)}: Observed occupancy rate of non-blocked days multiplied by 366 multiplied by median nightly rate

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Likely' revenue metric to be a better measure of actual market performance. We chart the distribution of the Likely ($REV_l$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties. Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year (\autoref{revdens}).  Actual short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not.  When looking at potential revenues, however, a great deal more may have the potential to produce additional revenue.  The high number of very low yield Airbnb properties in the Likely condition suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.]

<!-- ```{r plot_revdens, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=6,fig.height=3.5, fig.cap='\\label{revdens}Comparison of Revenue Distributions'} -->

<!--   ggplot(revdens.data, aes(x=revenue, fill=tenure, color=tenure)) + -->
<!--     geom_density(alpha=.5) + -->
<!--     scale_fill_manual(values=c(ltr.col, abb.col[2], str.col)) + -->
<!--     scale_color_manual(values=c(ltr.col, abb.col[2], str.col)) + -->
<!--     xlab('\nAnnual Revenue') + -->
<!--     scale_x_continuous(breaks=c(seq(0,60000, by=20000)), -->
<!--                        labels=c('$0', '$20k', '$40k', '$60k'))+ -->
<!--     theme(legend.position='none', -->
<!--           legend.title = element_blank(), -->
<!--           plot.title = element_text(hjust = 0.5), -->
<!--           axis.text.y = element_blank(), -->
<!--           axis.ticks.y = element_blank(), -->
<!--           axis.title.y=element_blank()) + -->
<!--     coord_cartesian(xlim=c(0, 70000)) + -->
<!--     facet_wrap(~tenure) -->

<!-- ### Comparison -->

<!-- In this section we compare net revenues between long- and short-term rental options.  As mentioned above, there are no direct matches between the datasets. In other words, we have not directly observed a property which has been in a long-term rental followed by short-term rental or vice versa.  We will, instead, make the comparisons between the two by imputing likely long-term rental revenues for the observed short-term properties and then comparing the measured net revenues of the short-term properties to their estimated imputed long-term revenues. We first focus on using the Actual measure of net revenue for the short-term properties as discussed above.  We save discussion and analysis of the Potential revenue for the sensitivity tests later in the study.  -->

<!-- We make our imputations via hedonic price models. We model the long-term market with the long-term data and then use this model to predict or impute the likely long-term weekly rental prices for the short-term properties. We use the following model specification: -->

<!-- \begin{equation} -->
<!--   \label{eq:test} -->
<!-- ln(RentLT_{i}) = \beta_{1} BedBath_{i} + \gamma_{2} Suburb_{i} + \tau_{3} Time_{i} + \epsilon_{i} -->
<!-- \end{equation} -->

<!-- where $ln(RentLT_{i})$ is the natural log of the weekly rent, *BedBath* is the bedroom/bathroom combination, *Suburb* is a fixed effect for the location of the property, *Time* is the month of rental and $\epsilon$ a random error term. -->
